---
output: github_document
bibliography: "references.bib"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# independenceWeights

<!-- badges: start -->
[![version](http://www.r-pkg.org/badges/version/independenceWeights)](https://cran.r-project.org/package=independenceWeights)
<!-- badges: end -->

The `independenceWeights` package constructs weights designed to minimize the weighted statistical dependence between a continuous exposure variable and a vector of confounder variables and implements the methods of @huling2021independence for doing so. In estimating a causal dose-response function, confounding bias is a function of the dependence between confounders and the exposure/treatment, so weights that minimize the dependence aim to mitigate confounding bias directly.

## Installation

You can install the released version of independenceWeights from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("independenceWeights")
```

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jaredhuling/independenceWeights")
```
## Example

This is a basic example which shows how to estimate and utilize the distance covariance optimal weights (DCOWs) of @huling2021independence:

```{r load, message=FALSE, warning=FALSE}
library(independenceWeights)
```

Simulate data with a continuous treatment that has a confounded relationship with a response. Data are simulated according to the simulation setup of @vegetabile2021nonparametric.

```{r sim}
simdat <- simulate_confounded_data(seed = 999, nobs = 500)
y <- simdat$data$Y ## response
A <- simdat$data$A ## treatment
X <- as.matrix(simdat$data[c("Z1", "Z2", "Z3", "Z4", "Z5")]) ## confounders

```

Now estimate weights to adjust for confounders using the distance covariance optimal weights (DCOWs), which aim to mitigate the dependence between $A$ and $X$:

```{r dcows}
dcows <- independence_weights(A, X)

dcows
```

Alternatively, information about any set of weights can be printed via `weighted_energy_stats()`.
The unweighted distance covariance is a measure of the unadjusted dependence between confounders
and treatment. The weighted dependence distance is a measure of this dependence *after* weighting.
The weighted energy distance between A and weighted A measures how well the marginal distribution
of A is preserved after weighting. Similarly for X. The effective sample size estimates the effective
sample size after weighting.

```{r wtstats}
weighted_energy_stats(A, X, dcows$weights)
```

We can assess traditional measures of balance such as marginal weighted correlations with `cobalt`:
```{r message=FALSE,warning=FALSE}
library(cobalt)
```

Even though the DCOWs aim to mitigate joint dependence between X and A, they result in very small
marginal weighted correlations between them as well. (changing `poly` to 10 below would reveal that
even up to 10th degree polynomials of covariates as well-decorrelated):
```{r}
bal.tab(A ~ Z1 + Z2 + Z3 + Z4 + Z5, data = data.frame(A = A, X),
        weights = list(DCOW = dcows$weights),
        un = TRUE, int = TRUE, poly = 1, stats = "cor")
```

We can also minimize the weighted dependence subject to the constraint that
marginal moments of X and A are decorrelated:
```{r dcowsdecor}
dcows_decor <- independence_weights(A, X,
                                    decorrelate_moments = TRUE)

dcows_decor
```

```{r}
bal.tab(A ~ Z1 + Z2 + Z3 + Z4 + Z5, data = data.frame(A = A, X),
        weights = list(DCOW = dcows$weights,
                       DCOW_decor = dcows_decor$weights),
        un = TRUE, int = TRUE, stats = "cor")
```

Now use the weights to estimate the causal average dose response function (ADRF)

```{r adrf, fig.height = 6}
## create grid
trt_vec <- seq(min(simdat$data$A), 50, length.out=500)

## estimate ADRF
adrf_hat <- weighted_kernel_est(A, y, dcows$weights, trt_vec)$est

## estimate naively without weights 
adrf_hat_unwtd <- weighted_kernel_est(A, y, rep(1, length(y)), trt_vec)$est

ylims <- c(-4.75, 4.75)
plot(x = simdat$data$A, y = simdat$data$Y, ylim = ylims, 
     xlim = c(0,50),
     xlab = "A", ylab = "Y")
## true ADRF
lines(x = trt_vec, y = simdat$true_adrf(trt_vec), col = "blue", lwd=2)
## estimated ADRF
lines(x = trt_vec, y = adrf_hat, col = "red", lwd=2)
## naive estimate
lines(x = trt_vec, y = adrf_hat_unwtd, col = "green", lwd=2)
legend("bottomleft", c("True ADRF", "Unweighted Est.", "DCOW-weighted Est."),
       col = c("blue", "green", "red"), lty = 1, lwd = 2)
```

## References